% SPDX-FileCopyrightText: 2023 SAP SE
%
% SPDX-License-Identifier: Apache-2.0
%
% This file is part of FEDEM - https://openfedem.org

\Section{Wind turbine modeling approaches}{wind-turbine-modeling-approaches}

There are alternative approaches to wind turbine modeling:

\begin{itemize}
\item{\sl NREL/FAST} --
  The FAST software is the first generation of wind turbine modeling tools.
  The software is developed as a part of a public research project at
  the National Renewable Energy Laboratory.
  The FAST software is open source, which is important in many research projects
  that seek to advance scientific knowledge within the area.
  The tool does, however, not have a graphical user interface and can be
  difficult to use correctly in a fast-pased commercial application.
\item{\sl GLGH/Bladed} --
  The Bladed software is the first generation of privately developed and
  commercially available wind turbine modeling tools.
  The tool basically adds a graphical user interface to the wind
  turbine modeling process. A set of dialog boxes are used to input
  options and data. The tool offers a straight-forward approach to the
  wind turbine modeling process that works well for typical wind turbine models.
\item{\sl Fedem WindPower} --
  We, respectfully, offer Fedem WindPower as a next generation wind turbine
  modeling environment. The Fedem software is designed and developed to be
  a high-end technology platform and an engineering framework for virtual
  testing of complex mechanical assemblies. It provides a complete set of
  features to create, solve and post-process wind turbine models in a
  3D graphical environment.
  We can offer our software as a complete replacement of prior software.
\end{itemize}

There are additional wind turbine modeling tools as well that are made
at universities, research institutions and elsewhere but they are in the
design line of first generation tools, for the most part.

Fedem WindPower is the only software that offers a high-end technology platform
and engineering framework for virtual testing of complex wind turbine models.
Fedem WindPower includes a large array of mechanism elements such as parts,
joints, springs, dampers, loads, control systems, and so on,
that can be added to (or modified in) the wind turbine model.
Each mechanism element has a large array of properties that offers fine detail
control over the model. The solvers produce results that offer an easy and
intuitive analysis approach with integrated graphs and tables.
Results can also be exported to other tools.
Never before have wind turbine modeling, solving and analysis been this easy,
flexible, extensible, comprehensible, precise, auditable and powerful!
