% SPDX-FileCopyrightText: 2023 SAP SE
%
% SPDX-License-Identifier: Apache-2.0
%
% This file is part of FEDEM - https://openfedem.org

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% FEDEM Theory Guide.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{thebibliography}{99}

\bibitem{AllmanTriPlate}
Allman, D.\ J.,
``A Simple Cubic Displacement Element for Plate Bending'',
\textit{Int.\ Jour.\ for Num.\ Meth.\ in Engrg.},
Vol.\ 10, pp.\ 263--281, 1976.

\bibitem{OIS:dring}
Sivertsen, O.\ I.,
``Large Displacement Finite Element Formulations of Elastic Mechanism Dynamics'',
\textit{Dr.\ Ing.\ Thesis},
Norwegian Institute of Technology, 1981.

\bibitem{BerganLarsenMollestad}
Bergan, P.\ G., Larsen, P.\ K.\ and Mollestad, E.,
``Svingning av konstruksjoner'',
Tapir forlag, 1981.

\bibitem{BerganFelippaTri}
Bergan, P.\ G.\ and Felippa, C.\ A.,
``A Triangular Membrane Element with Rotational Degrees of Freedom'',
\textit{Comput.\ Methods Appl.\ Mech.\ Engrg.},
Vol.\ 50, pp.\ 25--69, 1985.

\bibitem{RankinNourOmid}
Rankin, C.\ C.\ and Nour-Omid, B.,
``The Use of Projectors to Improve Finite Element performance'',
\textit{Computers and Structures},
Vol.\ 30, pp.\ 257--267, 1988.

\bibitem{RungeKutta}
Enright, W.\ H.\ et.\ al,
``Interpolants for Runge--Kutta Formulas'',
\textit{ACM Trans.\ on Math.\ Softw.},
Vol.\ 12, No.\ 3, pp.\ 193--218, 1986.

\bibitem{GearOsterby}
Gear, C.\ W.\ and O.\ {\O}sterby,
``Solving Ordinary Differential Equations with Discontinuities'',
\textit{ACM Trans. on Math. Softw.},
Vol.\ 10, No.\ 1, pp.\ 23--44, 1984.

\bibitem{HHT}
Hilber, H.\ M., Hughes T.\ J.\ R.\ and Taylor R.\ L.,
``Improved Numerical Dissipation for Time Integration Algorithms in Structural Dynamics'',
\textit{Earthquake Engineering and Structural Dynamics},
Vol.\ 5, pp.\ 283--292, 1977.

\bibitem{Iversen:STF48}
Iversen, T.,
``Parallel, Modular Integration for Dynamic Simulation of Industrial Processes'',
SINTEF report STF48 F86015; (in Norwegian), 1986.

\bibitem{Iversen:STFworkNote}
Iversen, T.,
``Multidisciplinary Simulation. Method, Software Structure and Documentation for the Control Part'',
SINTEF work note 88--63--K; (in Norwegian), 1988.

\bibitem{Newmark}
Newmark, N.\ M.,
``A Method of Computation for Structural Dynamics'',
\textit{J.\ Eng.\ Mech.\ Div.\ ACSE},
Vol.\ 85, EM3, 1959.

\bibitem{BH:phd}
Haugen, B.,
``Buckling and Stability Problems for Thin-Shell Structures using High-Performance Finite Elements'',
\textit{Ph.\ D.\ Dissertation},
University of Colorado, 1994.

\bibitem{GeradinRixen}
Geradin, M.\ and Rixen, D.,
``Mechanical Vibrations, Theory and Applications to Structural Dynamics'',
Wiley \& Sons Ltd., 1997.

\bibitem{HulbertChung}
Hulbert, G.\ M., Chung J.\ J.,
``A Time Integration Algorithm for Structural Dynamics with Improved Numerical Dissipation: The Generalized-$\alpha$ Method'',
\textit{J. Appl. Mech.},
Vol.\ 60, pp. 371--375, 1993.

\bibitem{Brincker}
Brincker, R.\ and Ventura, C.,
``Introduction to Operational Modal Analysis'',
Wiley, 2015.

\iftoggle{publicedition}{}{% The following is for the in-house edition only
\bibitem{Chakrabarti}
Chakrabarti, S.\ K.,
``Hydrodynamics of Offshore Structures'',
WIT Press, 1987.
} % End in-house edition only

\end{thebibliography}
