% SPDX-FileCopyrightText: 2023 SAP SE
%
% SPDX-License-Identifier: Apache-2.0
%
% This file is part of FEDEM - https://openfedem.org

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% FEDEM Theory Guide.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{Simulation Results}

\iftoggle{publicedition}{}{% The following is for the in-house edition only.

\section{Primary simulation variables}

The primary simulation variables are the supernode orientation and position,
the supernode translational and rotational velocity and acceleration,
and the current superelement transformation matrix.
Variables from the control simulation (the control variables), and the joint
variables are also primary variables in the simulation.

The orientation and position of the supernodes are represented as a $(3\times4)$
transformation matrix (see \eqsref{eq:posDef}{eqSC:Png}):
%
\begin{equation}
{\mf P}_{NG} \;=\; \left[ \begin{array}{*{20}c}
R_{11} & R_{12} & R_{13} & p_x \\
R_{21} & R_{22} & R_{23} & p_y \\
R_{31} & R_{32} & R_{33} & p_z \\
\end{array} \right] \;=\; \left[ \begin{array}{cccc}
{\mf i}_x & {\mf i}_y & {\mf i}_z & {\mf p}
\end{array}\right]
\label{eqK43:1}
\end{equation}
%
where ${\mf i}_x$, ${\mf i}_y$, and ${\mf i}_z$ are unit vectors in the $x$-,
$y$-, and $z$-directions, respectively, of the supernode coordinate system, and
the vector ${\mf p}$ represents the position of supernodes, all of which are
referred to in the global coordinate system.
The three unit vectors represent the direction cosine matrix, which uniquely
defines the orientation in space.

The supernode velocities are presented by two Cartesian vectors, one of which
represents the translational velocity, and one the rotational velocity relative
to the global coordinate system as shown below:
%
\begin{equation}
\left[\begin{array}{c} {\mf v_t} \\ {\mf v_r} \end{array}\right] \;=\;
\left[\begin{array}{cccccc}
v_{tx} & v_{ty} & v_{tz} & v_{rx} & v_{ry} & v_{rz}
\end{array}\right]^T
\label{eqK43:2}
\end{equation}
%
where
%
\begin{namelist}{$\quad{\mf v_r}$}
\item[$\quad{\mf v_t}$] represents the translational velocity vector, and
\item[$\quad{\mf v_r}$] represents the rotational velocity vector
(angular velocity).
\end{namelist}

The supernode accelerations are similarly presented as two Cartesian vectors
relative to the global coordinate system as shown below:
%
\begin{equation}
\left[\begin{array}{c} {\mf a_t} \\ {\mf a_r} \end{array}\right] \;=\;
\left[\begin{array}{cccccc}
a_{tx} & a_{ty} & a_{tz} & a_{rx} & a_{ry} & a_{rz}
\end{array}\right]^T
\label{eqK43:3}
\end{equation}
%
where
%
\begin{namelist}{$\quad{\mf a_r}$}
\item[$\quad{\mf a_t}$] represents the translational acceleration vector, and
\item[$\quad{\mf a_r}$] represents the rotational acceleration vector
(angular acceleration).
\end{namelist}

The magnitudes for the modal supernode displacement, velocity, and acceleration,
which represent fixed interface superelement modes of vibration
(described in Section~\ref{secCMS:FixeIntMode}), are presented in a similar way
to that of ordinary supernodes, with the exception that there are a variable
number of DOFs per supernode.
The variables for a modal supernode may be written:
%
\begin{equation}
\left[\begin{array}{ccc}
{\mf y} & \dot{\mf y} & \ddot{\mf y} \end{array}\right] \;=\;
\left[ \begin{array}{*{20}c}
y_1 & \dot{y}_1 & \ddot{y}_1 \\
y_2 & \dot{y}_2 & \ddot{y}_2 \\
\vdots & \vdots & \vdots \\
y_n & \dot{y}_n & \ddot{y}_n
\end{array} \right]
\label{eqK43:4}
\end{equation}
%
where
%
\begin{namelist}{$\quad\ddot{\mf y}$}
\item[$\quad{\mf y}$]      is the modal supernode displacement magnitude,
\item[$\quad\dot{\mf y}$]  is the modal supernode velocity magnitude, and
\item[$\quad\ddot{\mf y}$] is the modal supernode acceleration magnitude.
\end{namelist}

The superelement position and orientation are represented as a $(3\times4)$
transformation matrix (see \eqsref{eq:posDef}{eqK43:1}):
%
\begin{equation}
{\mf P}_{SG} \;=\; \left[ \begin{array}{*{20}c}
R_{11} & R_{12} & R_{13} & o_x \\
R_{21} & R_{22} & R_{23} & o_y \\
R_{31} & R_{32} & R_{33} & o_z \\
\end{array} \right] \;=\; \left[\begin{array}{cccc}
{\mf i}_x & {\mf i}_y & {\mf i}_z & {\mf o} \end{array}\right]
\label{eqK43:5}\end{equation}
%
where the unit vectors ${\mf i_x}$, ${\mf i_y}$ and ${\mf i_z}$ represent the
orientation of the superelement coordinate system, while ${\mf o}$ represents
the origin's position in the superelement coordinate system.

As previously stated, control variables are also primary simulation variables.
The control variables are defined as input, state, and algebraic variables.

\section{Derived simulation variables}

The following simulation variables are derived from the primary variables:
%
\begin{itemize}
\item   Spring variables
\item   Damper variables
\item   Forces and torques
\end{itemize}

A spring element is a uniaxial or a joint spring and has three variables:
The spring length/angle; the spring deflection from the stress-free position;
and the spring force/torque.
The spring length of a uniaxial spring is calculated from the position of the
two end nodes.
In case of joint springs, the spring length or angle refers to a joint variable.
Subtracting the specified stress-free length/angle of the spring from the
calculated length/angle of the spring gives the spring deflection.
In the general nonlinear case, the spring force/torque is found by integrating
the spring stiffness over the spring deflection.

A damper element is a uniaxial or joint damper and has three variables:
The damper length/angle; the damper relative velocity;
and the damper force/torque.
The damper length of uniaxial dampers is calculated from the position of
the two end nodes.
In case of joint dampers, the damper length or angle refers to a joint variable.
Damper velocity for uniaxial dampers is calculated from the end node velocities.
Denoting the global coordinates of the two end points of the damper as
%
\begin{equation}
{\mf p}_1 \;=\;
\left[ \begin{array}{*{20}c} x_1 \\ y_1 \\ z_1 \end{array} \right]
\quad\text{and}\quad
{\mf p}_2 \;=\;
\left[ \begin{array}{*{20}c} x_2 \\ y_2 \\ z_2 \end{array} \right]
\label{eqK43:6}
\end{equation}
%
the damper unit direction vector is calculated from
%
\begin{equation}
{\mf e}_{12} \;=\; \frac{{\mf p}_2 - {\mf p}_1}{\|{\mf p}_2 - {\mf p}_1\|}
\label{eqK43:7}
\end{equation}
%
The relative velocity for these two points may be calculated from
%
\begin{equation}
{\mf v}_{12} \;=\;
\left[ \begin{array}{*{20}c}
\dot{x}_1 \\ \dot{y}_1 \\ \dot{z}_1
\end{array} \right] -
\left[ \begin{array}{*{20}c}
\dot{x}_2 \\ \dot{y}_2 \\ \dot{z}_2
\end{array} \right]
\label{eqK43:8}
\end{equation}
%
and the damper velocity is then calculated from the scalar product
%
\begin{equation}
{\mf v}_D \;=\; {\mf e}_{12}^T {\mf v}_{12}
\label{eqK43:9}
\end{equation}
%
In the case of joint dampers, the damper velocity refers directly to a joint
variable velocity.
The damper force/torque is found by multiplying the damping coefficient by the
damper velocity.

For calculation of superelement node forces and torques, refer to the stiffness
relation of \eqnref{eqSC:forceLocal}, see Section~\ref{sec:SupElIntForce}.
The forces are transformed to global direction and added into the system vector.
This system vector is then used to present the nodal forces and torques in
relation to a global coordinate system.
Joint forces and torques are found on the corresponding master or slave DOFs
of the joint, and based on the superelement forces from deformation,
\eqnref{eqSC:forceLocal}, combined with the contributions from inertia
and damping, we get:
%
\begin{equation}
\tilde{\mf f}_e \;=\;
\tilde{\mf k}_e \tilde{\mf v}_d +
\tilde{\mf c}_e \dot{\tilde{\mf v}} +
\tilde{\mf m}_e \ddot{\tilde{\mf v}}
\label{eqSR:TotalSupElForce}
\end{equation}

\section{Substructure retracking}

The co-rotated superelement formulation developed in
Chapter~\ref{chap:SupElCorot} shows the derivation of the superelement
deformation vector for an external node, \eqnref{eqSC:nodeDefDisp}.
The modal coordinates ${\mf y}$ are primary integration variables;
and the superelement deformation vectors for all external nodes and modal
coordinates are given by \eqnref{eqSC:supelDefDisp}.

The inverse CMS transformation gives
%
\begin{equation}
\label{eqK43:10}
\eqalign{
{\mf v}_i \;=\; & {\mf v}_i^i + {\mf v}_i^e \cr
            =\; & {\mf B} {\mf v}_e + {\tf \Phi}{\mf y}
}
\end{equation}
%
and the expanded substructure displacement vector may be written
%
\begin{equation}
{\mf v} \;=\;
\left[ \begin{array}{c} {\mf v}_e \\ {\mf v}_i \end{array} \right] \;=\;
{\mf H} \left[ \begin{array}{c} {\mf v}_e \\ {\mf y} \end{array} \right]
\label{eqK43:11}
\end{equation}
%
(see \eqnref{eqCMS:Hdef}).

Nodal displacements for substructure primary element $j$ may now be extracted
from the substructure displacement vector:
%
\begin{equation}
{\mf u}_j \;=\; {\mf A}_j {\mf v}
\label{eqK43:13}
\end{equation}
%
where
%
\begin{namelist}{${\mf u}_j$}
\item[${\mf u}_j$] represents the displacement vector of primary element $j$
in the substructure,
\item[${\mf A}_j$] is the incidence matrix that represents the topology
for primary element $j$ in the substructure, and
\item[${\mf v}$] is the substructure displacement vector of \eqnref{eqK43:10}.
\end{namelist}

\section{Finite element stress analysis}

The strain field within a finite element is expressed by the nodal DOFs
${\mf u}$ and a set of interpolation polynomials.
For instance, for a membrane finite element:
%
\begin{equation}
{\mf\varepsilon}(x,y) \;=\; \bar{\mf B}(x,y)\,{\mf u}
\label{eqK43:17}
\end{equation}
%
where
%
\begin{namelist}{$\quad\bar{\mf B}$}
\item[$\quad{\tf\varepsilon}$] is the element strain vector, and
\item[$\quad\bar{\mf B}$]      is the strain interpolation matrix,
\end{namelist}
%
The stress is then according to Hooke's Law:
%
\begin{equation}
{\tf\sigma} \;=\; {\mf C} {\tf\varepsilon}
\label{eqK43:18}
\end{equation}
%
where
%
\begin{namelist}{$\quad{\mf E}$}
\item[$\quad{\tf\sigma}$] is the element stress vector, and
\item[$\quad{\mf C}$] is the constitutive matrix.
\end{namelist}
%
For further details, refer to textbooks on the finite element method.

\input{strain_gages}
} % End in-house edition only
\input{fatigue_analysis}
\iftoggle{publicedition}{}{% The following is for the in-house edition only.
\input{eigenvalue_problem}
} % End in-house edition only
\input{energy_calc}
